Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques

### 2018-01-11 paper/gloaguenNmfTraffic.tex
done with intro

With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict an estimation of the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}. These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques. They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}. In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day. Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.

Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels. DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations. It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}. In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.

If sensor networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}. Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn \dots) that can overlap. Consequently, the traffic sound level estimation based on measurements is not a trivial task.
Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two-step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

The main issue in the detection or recognition tasks is the overlap of environmental sounds. Although near major roads or ring roads, traffic is predominant, there are many places where it overlaps with other sound sources that contribute significantly to the overall sound levels. Succeed to recognize two simultaneous sound sources can be very complicated. Socoro et al. propose to bypasses this issue with an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level. If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.

Instead of using a detection method to estimate the traffic presence, we propose in this paper to follow the blind source separation paradigm. That is, separating the contribution of the traffic from the other sources within a polyphonic scene. \\

One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}. The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations. This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise. In an urban environment context, this method is used with acoustic microphone arrays and beamforming \cite{saruwatari2003blind}. These approaches consider a set of microphones and allows the localization \cite{mennitt2010multiple} or the detection \cite{leiba2017large} of sound sources from the phase shift and the distance between the microphones. However, one need a lot of microphones. Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.

A more convenient method for monophonic signal is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices. It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane. For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.

For the sound source separation, it has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012}. After having performed NMF on simulated audio files, they realized a source separation in two steps by separating the sound background from the events first and then by separating the events between them using MFCC and time-variant gain. A $k$-mean clustering allows the separation of the $k$ sound sources. This approach, compare to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.  Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.

Consequently, we propose in this paper to use different frameworks of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event. To validate this approach, we also consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples. The use of simulated sound scenes is necessary as it offers a full control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would hardly be extracted from a recording of an urban scene.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
\caption{Bloc diagram of the source separation method}
\end{figure*}

The remaining of the paper is organized as follows. Section \ref{part:nmf} details the technical aspects of NMF. Section \ref{part:protocol} described on the design of the environmental sound scene corpus and the experimental protocol setup. Then Section \ref{part:results} shows and discusses the results obtained during the parametric study.

>>>>

With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict an estimation of the number of city dwellers exposed to high noise levels. They are used to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}. These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques. They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} \ml{un peu hors propos on améliore rien la dessus : or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}. plutot parler du cote sparse des données d'entrées}. In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day. Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.

Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters ...), including the assessment of urban noise levels. DYNAMAP project \cite{dynamap_2016} studied the deployment and feasibility of such installations. It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}. In a similar way, but reduced to few neighborhoods \ml{with a more dense topology ? parler aussi de sonyc}, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.

If sensor networks could improve road traffic noise estimates compared with simulated maps, the issue of the correct estimation of the traffic sound level from acoustic measurements is still unsolved \cite{Mioduszewski}. Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes \dots) that overlap. Consequently, the traffic sound level estimation based on measurements is not a trivial task.
Many recent works have focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two-step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

The main issue in the detection or recognition tasks is the overlap of environmental sounds. Although near major roads or ring roads, traffic is predominant, there are many places where it overlaps with other sound sources that contribute significantly to the overall sound levels. To circumvent this issue, Socoro et al. propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017} which consist in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component. Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.  \ml{arrete :) If the detection of the road traffic noise is good, the detection of these anomalous noise events stay a complicated task \cite{stowell2015detection}.}

An alternative approach that we will foolow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level. That is, separating the contribution of the traffic from the other sources within a polyphonic scene. One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\

ml{Supprimer : One of the first to do so is the Independent Component Analysis \cite{comon_independent_1994}. The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations. This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise. In an urban environment context, this method is used with acoustic microphone arrays and beamforming \cite{saruwatari2003blind}. These approaches consider a set of microphones and allows the localization \cite{mennitt2010multiple} or the detection \cite{leiba2017large} of sound sources from the phase shift and the distance between the microphones. However, one need a lot of microphones. Spread multiples microphones arrays in cities is then expensive (even with low cost microphones) and is time-consuming for calibration and maintenance.}

A convenient method for monophonic signal decomposition is the Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}. When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}. In the audio processing domain, the NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane. \ml{enlever : For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded.}

Closer to our application scenario, NMF has been considered by Innami and Kasai \cite{satoshi_innami_nmf-based_2012}. After having performed NMF on simulated audio files, they perform a source separation based on NMF in two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.

\ml{ enlever : A $k$-mean clustering allows the separation of the $k$ sound sources. This approach, compared to a simple MFCC clustering, improves the sound source separation but the authors admit their approach need the assumption that the number of sound sources is known which is unrealistic in practical case.  Furthermore, in order to estimate the traffic sound level $\tilde{L}_{p,traffic}$, their approach might be difficult to be adapted it as traffic component can be both sound background and event.}

We study in this paper different flavors of Non-Negative Matrix Factorization (NMF) where traffic component is considered in its entirely whether it is a sound background or event. We demonstrate the supervised NMF and semi supervised NMF approaches have some interests but fail to give satifactory results for the application at hand. We thus introduce another NMF scheme called thresholded initialized NMF that make good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio (TIR).

To perform the numerical experiments, we consider a corpus of simulated sound scenes created from a built-up sound database composed of a high number of diverse sound samples. The use of simulated sound scenes is mandatory for rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$) which would be difficult to extract from a recording of an urban scene.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
\caption{Block diagram of the source separation method. \ml{toujours citer une figure}}
\label{fig:diagram}
\end{figure*}

The remaining of the paper is organized as follows. Section \ref{part:nmf} details the technical aspects of NMF and describes the 3 approaches considered in this paper to achieve the task at hand. Section \ref{part:protocol} describes the environmental sound scene corpus and the experimental protocol setup. Section \ref{part:results} then describes and discusses the outcomes of the numerical experiments.
